---
title: "More Visualizations"
author: 
- "Fynn Hayton-Ruffner"
- "Shekinah Ninziza"
- "Kinglee Tham"
---

```{r}
#| include: false


library(ggplot2)
library(tidytext)
library(tidyverse)
library(RColorBrewer)
library(ggthemes)
library(textdata)
library(lubridate)
library(viridis)
library(ggiraph)

```


```{r}
#| label: fig-word_tfidf


load('./data/top_word_tfidf.Rdata')
load('./data/top_bigram_tfidf.Rdata')

top_word_tfidf |>
  ggplot(aes(x = reorder_within(word, tf_idf, subreddit), y = tf_idf, fill = tf_idf)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  facet_wrap(~subreddit, ncol = 2, scales = "free") +
  scale_x_reordered() +
  labs(x = NULL, 
       y = "TF-IDF",
       title = "Top 10 words by Tf-Idf for Each Subreddit") +
  scale_fill_viridis('magma')


```

Amherst College 

The most significant TF-IDF term is "amherst's", emphasizing the strong connection that posts on this subreddit have with local institutional discussions. Other significant terms such as "umass," "statistics," and "sem" (presumably an abbreviation for "seminar") indicate that academic conversations are crucial. Furthermore, terms such as "soccer" and "paper" indicate both extracurricular activities and subjects related to coursework. This combination of academic and student life language reflects a comprehensive involvement with both the intellectual and daily aspects of Amherst. 

Middlebury College 

The notable term in this context is "midd"( an abbreviation for Middlebury recognized by its community), its elevated TF-IDF score suggests significant internal usage of the word which might be rare out of the school. Additional terms such as "Vermont," "MIIS" (denoting the Middlebury Institute of International Studies), and "immersion" suggest the institution's language offerings and regional character. The significance of these phrases indicates a subreddit focused on regional culture, international academic initiatives, and university branding. 

Williams College 

The subreddit for Williams is primarily focused on "williamstown", which is the town where Williams is so they probably talk more its location. Words such as "loci" (probably an abbreviation for "Letter of Continued Interest"), "supplement," and "previews" indicate a high level of involvement with the admissions procedure. The existence of "ephs" (the mascot of Williams), along with mentions of peer schools such as "dartmouth", highlights a mixture of institutional pride and competitive rivalry. The subreddit seems to be very busy throughout the admissions period and centered on unique community slang and customs. 



```{r}
#| label: fig-bigram_tfidf


 top_bigram_tfidf |>
  ggplot(aes(x = reorder_within(bigram, tf_idf, subreddit), y = tf_idf, fill = tf_idf)) +
  geom_col() +
  coord_flip() +
  theme(legend.position = "none") +
  facet_wrap(~subreddit, ncol = 2, scales = "free") +
  scale_x_reordered() +
  labs(x = NULL, 
       y = "TF-IDF",
       title = "Top 10 Bigrams by Tf-Idf for Each Subreddit") +
  scale_fill_viridis('magma')

```



How to use RedditExtractoR [@Ant_2023]


Each page can have it's own list of references based on what you cite. For example, did you know that good graphs "have a clear purpose, show the data clearly, and make the message obvious"? [@vandem2019].